---
title: "ASMA Reference Table"
author: "RQ"
date: "11 February 2019"
output: html_document
---

This RMarkdown script documents the data preparatory action to trim the ASMA Reference Table for this study.

It broadly 

* reads in the ASMA REF TABLE
* subsets for the study airports, i.e. EGLL, EIDW, and LSZH
* corrects for the sector spanning the northerly bearing of 360
* writes out the reference table for this study


## Extracting the ASMA Reference Table

The ASMA Reference Table is extracted from PRISME with the following SQL code.
Note: This is done by hand, i.e. files saved locally, as the ROracle interface could not be used.
Conceptually, the reference table will be made available to the public.


## Reading in and Subsetting the ASMA Reference Table 

```{r, message=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(purrr)
```
```{r, message=FALSE}
study_apts <- c("EGLL", "EIDW", "LSZH")

asma_refs <- readr::read_csv2("../data/ASMA_REF_20190201.csv") 

asma_refs <- asma_refs %>%
  filter(AIRPORT %in% study_apts & ASMA_RADIUS == 40) %>% 
  group_by(AIRPORT) %>%
  filter(VERSION == max(VERSION)) 

asma_refs <- asma_refs %>% group_by(AIRPORT, VERSION, ASMA_RADIUS) %>%
  tidyr::nest()
```

  
  
```{r}
# utility function checks if Bearing spans over North, i.e. 360.
# if condition holds, the range will be broken down into 2 sets
# a.) bearing to 360, and b.) 0 to end of span.
# all other values can stay the same, e.g. ASMA Sector, Unimpeded Time
# the changes rows will be merged to the data set
check_north_span <- function(ds){
  new_rows <- ds %>% group_by(AC_CLASS, RUNWAY) %>% 
    filter(FROM_BEARING == max(FROM_BEARING), !is.na(UNIMPEDED_TIME)) %>% 
    mutate(NORTH = TO_BEARING < FROM_BEARING ) %>% filter( NORTH)
  
  if(nrow(new_rows) >= 0){ message("correct for spanning north!")
    tmp1 <- new_rows %>%
      mutate(TO_BEARING = 360)
    tmp2 <- new_rows %>%
      mutate(FROM_BEARING = 0)
    
    ds <- anti_join(ds, new_rows) %>% bind_rows(tmp1) %>% bind_rows(tmp2)
  }else{
    message("ASMA sectors not spanning North!")
  }
  ds <- ds %>% select(-NORTH)
  return(ds)
}

# test for one case: tmp <- asma_refs$data[[1]] %>% check_north_span()
# apply to multiple cases
asma_refs <- asma_refs %>% mutate(REF = purrr::map(data, ~ check_north_span(.x)))

refs <- asma_refs %>% select(AIRPORT, VERSION, ASMA_RADIUS, REF) %>%
  tidyr::unnest() %>%
  select(AIRPORT, VERSION, ASMA_RADIUS, AC_CLASS, RUNWAY, ASMA_SECTOR, FROM_BEARING, TO_BEARING, UNIMPEDED_TIME) %>%
  arrange(AIRPORT, VERSION, ASMA_RADIUS, AC_CLASS, RUNWAY, ASMA_SECTOR)
```

Write out reference table for ATM Seminar paper.

```{r}
if(!file.exists("../data/ASMA_REF.csv")){
  message("Writing out trimmed refernce table for paper!")
  readr::write_csv(refs, "../data/ASMA_REF.csv")
}else{
  warning("Reference Table for paper already exists - stopping to write out!")
}
```


